---
title: "EDAV Fall 2020 PSet 4"
author: "null"
output:
  pdf_document: default
  html_document: default
---

This assignment is designed to help you get started on the final project. Be sure to review the final project instructions (https://edav.info/project.html).
    
### 1. The Team

[8 points]

a) Who's on the team? (Include names and UNIs)
\
 names:       UNI:  \
 Yue Wang     yw3576  \
 Haoyu Liu    hl3299  \
 James Ding   jd3703
 
b) How do you plan to divide up the work? (Grading is on a group basis. The point of asking is to encourage you to think about this and design a plan.)
\
  Right now, we came up with three research questions, but there will be more. Our idea is to come up with more questions related to a larger topic, and each of us will focus on one or two research questions, create graphs and provide comments. Each of us will add on to the others, communicates progress, and finally organize everything together. 

c) What is your plan for managing the git/GitHub workflow? That is, who will merge pull requests? How will you attempt to avoid conflicts? How will you communicate?
We will make sure everyone do it in our own branch, and merge everything together.. Yue Wang will be the person whoe will merge pull requests. In  order to avoid conflicts, we will communicate with each other once we have updated something through email to make sure they use the most updated file when folk it.\

### 2. The Questions

[10 points]

List three questions that you hope you will be able to answer from your research. (It's ok if these change as your work progresses.)
  \
a)  What are some factors that effect employment?
  \ 
b)  How do developed countries compare with developing countries in terms of job related areas?
  \  
c)  How do population structures vary across different conutries?
  
### 3. GitHub repo

[8 points]

a) Set up your final project repository following the EDAVproject template. Provide the link to the repo.
  \
https://github.com/SUN82564/5702-Final-Project

b) Make sure that all team members have write access to the repository and have practiced making contributions. Provide a link to the contributors page of your repository showing that all team members have made contributions to the repo (Note that we do not have the ability to see who has write access, only contributors):
  \
https://github.com/SUN82564/5702-Final-Project/graphs/contributors

### 4. Data Sources

[8 points]

Write a draft of the [Data Sources chapter](https://edav.info/project#report-format).
  \
The data source we use will be from the world bank database, and you can easily access the data from this link: https://databank.worldbank.org/source/jobs# . Yue will be responsible for collecting the data, and the process is relatively simple. One can go to the link and customize the data he/she wants from the interactive selection area. For out final project, we will be using the Jobs database, selecting 4 to 5 countries including Japan, China, United State, and South Africa, and all variables/series they provided. We will set the time range to be the most up to date 4-5 years to analyze. We will set the aggregation rule to be average of data available for each time period because we want to compare the overall situations across all countries. 
  \
Some basic information about the dataset:
  \
The job database provides 166 series/variables that we can use. Currently, we chose 5 countries and world data from most recent available 5 years, which yields 793*5 records. We work from this comprehensive dataset and select several variables that could help answer our research questions. We will also avoid using those that have a lot of missing values. The data types include categorical variables like countries, time series data, some variables with factor levels, and numerical variables. 
  \
Values for some variables are missing across all countries selected across all years, for example, the literacy rate, and values for variables like primary completion rate are missing for several years. We will further analyze the missing data to decide which ones could be used or not. Furthermore, the dataset is rather raw and will need to be reorganized because they break down into a very detailed way. For instance, many variables/series they provide should actually be the levels of the same variable. We will need to work on organize the data into a more usable form. 


### 5. Data Transformation

[8 points]

Write a draft of the [Data Transformation chapter](https://edav.info/project#report-format)

```{r message=FALSE, warning=FALSE}
library(readr)
data <- read.csv("Data_Extract_From_Jobs.csv",na.strings='..')
cleaned_data <- na.omit(data)
```
  \
Fortunately, the data which the website provided is in xlsx and provided with a detailed explanation about the variables inside the data. It could be easily converted to csv file which means that reading by Rstudio is straightforward. After finishing loading the data, we found that the data is already tidy. By now, we only need to change the missing value into *NA* and remove those *NA* values. \

After starting the project, we probably need to do more data transformation in order to display the interactive visualization and show complex information in d3.
  \


### 6. Missing Values

[8 points]

Write a draft of the [Missing Values chapter](https://edav.info/project#report-format)

```{r error=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(naniar)
data = read.csv("Data_Extract_From_Jobs.csv")
data <- data %>% replace_with_na_all(condition = ~.x == "..")
data <- data %>% filter(row_number() <= n() - 5)
```

Since countries have different priorities when gethering and reporting employment data, there are a lot of missing values in the dataset.

```{r}
colSums(is.na(data))
```
Understandably, the dataset does not have missing data regarding the country's name or the names of the factor (series) being measured. Furthermore, it seems that yearly the amount of missing data is around 210. Looking at the data, much of the missing data is not reported for all 5 years, which results in the consistent number of missing values. However, in 2013 the amount of missing data seemed to be lower at 195, and in 2016 the amount of missing data was much higher at 275.

```{r}
data$missing <- rowSums(is.na(data))
ggplot(data, aes(x = missing)) +
  geom_bar()
```

As shown above, for the vast majority of the rows, there are 0 missing values. This makes sense, as usually if the country reports the data in one year it'll usually continue to report the data for every year.  The next most common value seems to be 5, which means that the country simply does not report that data. The interesting values are 1 - 4, which means that the data was reported for some but not all of the years.

```{r}
ggplot(data, aes(x = missing)) +
  geom_bar() +
  facet_wrap(~ Country.Code)
```

A similar pattern seems to hold for the individual countries, where 0 is the most common number of missing values. However, for India, Japan, and South Africa, the second most common number of missing values are 4, 1, and 2, respectively. Since these are the countries with "unusual" distributions of missing values, it seems likely that these are the countries that cause the unusual number of missing values in year 2013 and 2016.

```{r}
Japan_missing <- data %>% filter(Country.Code == "JPN") %>% select(X2012..YR2012., X2013..YR2013., X2014..YR2014., X2015..YR2015., X2016..YR2016.) 
colSums(is.na(Japan_missing))
```
Japan reported slightly less missing values than average in 2013, which lines up with the global trend. However, the big difference is that Japan reported significantly more missing values in 2016. This probably contributed to the high number of missing data in 2016

```{r}
India_missing <- data %>% filter(Country.Code == "IND") %>% select(X2012..YR2012., X2013..YR2013., X2014..YR2014., X2015..YR2015., X2016..YR2016.) 
colSums(is.na(India_missing))
```

It seems that India is reporting more missing values as time goes on, which slightly contributes to the unusual values in 2013 and 2016.

```{r}
SA_missing <- data %>% filter(Country.Code == "ZAF") %>% select(X2012..YR2012., X2013..YR2013., X2014..YR2014., X2015..YR2015., X2016..YR2016.) 
colSums(is.na(SA_missing))
```

Contrary to the global trend, South Africa actually reported its second most missing values in 2013. It also reported its most missing values in 2016 (almost double its median), which certainly impacted the total number of missing values. Also, out of the 5 countries we are looking at, South Africa reports the least amount of missing values.

```{r}
China_missing <- data %>% filter(Country.Code == "CHN") %>% select(X2012..YR2012., X2013..YR2013., X2014..YR2014., X2015..YR2015., X2016..YR2016.) 
colSums(is.na(China_missing))
```

As expected, China was fairly consistent in its missing values, with slightly less in 2013 and slightly more in 2016. Out of the 5 countries we are looking at, China seems to have the most missing values.

```{r}
US_missing <- data %>% filter(Country.Code == "USA") %>% select(X2012..YR2012., X2013..YR2013., X2014..YR2014., X2015..YR2015., X2016..YR2016.) 
colSums(is.na(US_missing))
```
The United States also reports a large number of missing values. Like other countries, it reports its least amount of missing values in 2013 and most in 2016. However, its second largets number of missing values occurs in 2012.

```{r}
world_missing <- data %>% filter(Country.Code == "WLD") %>% select(X2012..YR2012., X2013..YR2013., X2014..YR2014., X2015..YR2015., X2016..YR2016.) 
colSums(is.na(world_missing))
```

In general, the world wide statistics are fairly consistent, with slightly more missing at 2016. It seems that the increase in missing values in 2016 is a world wide phenomenon for some reason.
